\section*{Analysis}
\label{empirics}

\subsection*{Estimation Method}

To model aid flows using our directed-dyadic panel dataset, we utilize a hierarchical model. To implement this we nest receivers within senders and senders within years. We include random intercepts in our model for every sender and year. The results of this analyis are shown below in table \ref{tab:mainResults}.

\begin{table}[!htbp] \centering 
\begin{tabular}{@{\extracolsep{5pt}}lc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ 
\cline{2-2} 
\\[-1.8ex] & Ln(Aid$_{sender-receiver, t}$) \\ 
\hline \\[-1.8ex] 
 Pol. Rel.$_{s-r, t-1}$ & 0.059$^{***}$ \\ 
  & (0.009) \\ 
  & \\ 
 Mil. Rel.$_{s-r, t-1}$ & $-$0.004$^{**}$ \\ 
  & (0.002) \\ 
  & \\ 
Colony$_{s-r, t-1}$ & 1.822$^{***}$ \\ 
  & (0.058) \\ 
  & \\ 
 Polity$_{r, t-1}$ & 0.015$^{***}$ \\ 
  & (0.002) \\ 
  & \\ 
 Ln(GDP Capita)$_{r, t-1}$ & $-$0.580$^{***}$ \\ 
  & (0.017) \\ 
  & \\ 
 Life Expect$_{r, t-1}$ & 0.017$^{***}$ \\ 
  & (0.002) \\ 
  & \\ 
 No. Disasters$_{r, t-1}$ & 0.168$^{***}$ \\ 
  & (0.004) \\ 
  & \\ 
 Constant & 17.463$^{***}$ \\ 
  & (0.107) \\ 
  & \\ 
\hline \\[-1.8ex] 
Observations & 34,486 \\ 
RMSE & 2.09  \\ 
\hline 
\hline \\[-1.8ex] 
  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
  \caption{Hierarchical regression results on logged aid flows between senders and receivers from 1975 to 2005.} 
  \label{tab:mainResults} 
\end{table} 

Our strategic interest variables are shown in the first two rows. The results for the political strategic relationship variable align with the extant literature, we can see that countries are likely to send greater levels of aid to those with whom they have strong political relationships. The same does not hold for strong military relationships, the relationship with this variable and aid is actually slightly negative. We also see that countries are more likely to send aid to former colonies, a finding which also has received support in previous literature. 

Next we turn to particular characteristics of receiver countries that are associated with higher levels of aid flows. Our analysis finds that countries which are poorer, more democratic, or that have recently faced a natural disaster receive higher levels of aid on average. Lower levels of life expectancy, however, are not associated with greater levels of aid. 

